---
title: "Practical Machine Learning Project: Exercise Type Prediction"
author: "Alex Gascoigne"
email: "agascoig@yahoo.com"
date: "January 30, 2016"
output: html_document
---

## Summary

This project involves exercise type prediction using accelerometer data on belt, forearm, arm,
and dumbell of 6 participants.  They performed barebell lifts correctly and incorrectly in
5 different ways.  The goal of the project is to predict the manner in which they did the
exercise, the "classe" variable.

The original source of the data and more information is available at:

http://groupware.les.inf.puc-rio.br/har

## Data Analysis

We load the data into R to train a random forest predictor, after subsetting the relevant
acceleration fields.  The training data is partitioned into **training** and **testing** sets. Then
K-fold (K=5) cross-validation is used with 3 repetitions.  This splits the data into five random partitions: four
parts are used for training, and one part validation.

No preprocessing of the data (centering, scaling) was done.

```{r}
library(caret)
set.seed(12345)
raw_train_data<-read.csv('training.csv')
raw_test_data<-read.csv('testing.csv')

vars<-c(
"total_accel_belt",     "accel_belt_x",         "accel_belt_y",         "accel_belt_z",        
"total_accel_arm",      "accel_arm_x",          "accel_arm_y",          "accel_arm_z",        
"total_accel_dumbbell", "accel_dumbbell_x",     "accel_dumbbell_y",     "accel_dumbbell_z",   
"total_accel_forearm",  "accel_forearm_x",      "accel_forearm_y",      "accel_forearm_z"
)

train_data<-subset(raw_train_data,select=c("classe",vars))
test_data<-subset(raw_test_data,select=vars)

inTrain = createDataPartition(train_data$classe, p = 0.7)[[1]]

training = train_data[ inTrain,]
testing = train_data[-inTrain,]

ctrl <- trainControl(method="repeatedcv",repeats=3,number=5)

print(date())
fit1<-train(classe~.,method="rf",data=training, trControl=ctrl,allowParallel=TRUE)
print(date())
fit1

pred<-predict(fit1,testing)
accuracy<-sum(pred==testing$classe)/dim(testing)[[1]]
print(accuracy)
```

The accuracy achieved by this predictor was `r accuracy` using 4,904 data points withheld from the training.

Finally, we predict the values for the original test set of 20 items:

```{r}
test_pred<-predict(fit1,test_data)
print(test_pred)
```

## Conclusion

Relatively high accuracy (94%) was achieved using a random forest predictor.